---
title: "MMR Example"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{example}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(mmr)
library(dplyr)
library(ggplot2)

theme_set(theme_bw())
```

# Introduction

The `mmr` package is designed to be used in preclinical AD cohorts. It allows to compute two different measure of awareness: the *Meta-Memory Ratio* (MMR), and the *Awareness of Cognitive Decline Categorization* (ACDC). Both of these scores are based on the Participant-Performance Discrepancy, i.e. the difference between a participant's subjective appraisal of her/his cognition, and her/his objective performance.

The MMR is numeric continuous measure, centered around 0 which corresponds to a good match between subjective and objective performance. A positive score indicates a SCD, and a negative stands for a lack of awareness.

The ACDC, based on the same inputs, categorized the participant as:

  * *Normal*: Good objective and subjective performances.
  * *Amnestic*: Low objective and subjective performances.
  * *SCD*: Low subjective but normal objective performance.
  * *LowAwareness*: normal subjective but low subjective performance.
  
Both of the measures are designed to be used in multiple preclincal AD samples. Here's a minimal example to show how to use the functions.

# Example

The `mmr` package contains a `data.frame` example 

```{r load_data}
data("testdf")
```

```{r head}
head(df.mmr)
```

```{r}
str(df.mmr)
```

Detail of data:

  * *IDNUM*: Participant Identification Number
  * *COHORT*: Cohort Name (here "Example")
  * *VISIT*: Visit ID (e.g. Baseline, M6, M12, etc.)
  * *AGE*: Age in years
  * *SEX*: Participant Gender
  * *EDUC*: Participant years of education
  * *FDG*: Meta-ROI of FDG-PET values from AD ROI
  * *AV45*: Mean of global amyloid-PET
  * *Mem1:6*: Memory performances
  * *SelfMemComp*: Memory Complaint Participant's Score
  * *SelfComp*: Participant's Complaint Score
  * *ExtMemComp*: Memory Complaint Relative's Score
  * *ExtComp*: Relative's Complaint Score


## Wrangle

The only prerequisite of the `mmr` functions to work is that the `data.frame` must follow some standards:
  
  1. Tidy Data: the dataset must be tidy with 1 row for every observations, and 1 column for every variable.
  1. Demographics: the demographics variables must be named as presented below.


### Demographics

First, we have to change the demographic variables names to fit the `mmr` package standard.

```{r rename_demogs}
df.mmr <- df.mmr %>%
  dplyr::rename(ID = IDNUM, # Participant ID,
                Gender = SEX, # Gender Variable Name,
                Education = EDUC, # Education Variable Name, 
                Age = AGE, # Age Variable Name
                Cohort = COHORT # Cohort Name (here "Example")
                ) 
```

### Vectors

Secondly, we have to define some vectors that will be usefull for the analysis.

```{r vectors}
demographics <- c("ID", "Gender", "Education", "Age")
markers <- c("AV45", "FDG")
complaint <- c("SelfMemComp")
memory <- c("Mem1", "Mem2", "Mem3", "Mem4", "Mem5", "Mem6")
```

### Distribution

The third step is to check the distribution of the scores. Then, we will have to decide wether we will treat the variables with linear or binomial regressions.

```{r check_dist}
df.mmr %>% dplyr::select(complaint, memory) %>%
  tidyr::gather(key = "Measure", value = "Score", c(complaint, memory)) %>%
  ggplot(aes(x = Score)) + geom_histogram(binwidth = 0.1) +
  facet_wrap(~ Measure, scales = 'free')
```

For the example, we will treat the `Mem4`and `Mem6` variables *via* binomial regressions. In our example, the `Mem4` variable was quoted on 75, and the `Mem6` on 25. All other variables will be treated with linear regressions.

In order to make thing as simple as possible, we will save this information in R objects.

```{r def_dist}
linear <- c("SelfMemComp", "Mem1", "Mem2", "Mem3", "Mem5")
binomial <- tibble::tibble(tests = c("Mem4", "Mem6"), # Measures that will be considered as binomial
                   max = c(75, 25)) # Max scores of above measures (in the same order)
```

Now we have all we need to build the awareness scores for our sample.

## The `mmr_build` function

```{r mmr_build}
df.mmr <- mmr_build(mem = memory,
                    comp = complaint,
                    vect.linear = linear,
                    vect.binomial = binomial,
                    df = df.mmr)
```

And that's it! As you can see below, we have our awareness measures. The below table represent a minimal information table sufficient to export if you want to add other cohorts or share data with others.

```{r mmr_min}
df.mmr %>% dplyr::select(demographics, Cohort, markers, Memory, Complaint, MMR, ACDC) %>% head()
```
